---
title: "Getting Started with the AppEEARS API (Area Request)"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    theme: paper
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dirname(rprojroot::find_rstudio_root_file()))

```

**This tutorial demonstrates how to use R to connect to the AppEEARS API**

The Application for Extracting and Exploring Analysis Ready Samples ([AppEEARS](https://appeears.earthdatacloud.nasa.gov/)) offers a simple and efficient way to access and transform [geospatial data](https://appeears.earthdatacloud.nasa.gov/products) from a variety of federal data archives in an easy-to-use web application interface. AppEEARS enables users to subset geospatial data spatially, temporally, and by band/layer for point and area samples. AppEEARS returns not only the requested data, but also the associated quality values, and offers interactive visualizations with summary statistics in the web interface. The [AppEEARS API](https://appeears.earthdatacloud.nasa.gov/api/) offers users **programmatic access** to all features available in AppEEARS, with the exception of visualizations. The API features are demonstrated in this tutorial.

## Example: Submit an area request using a U.S. National Park boundary as the region of interest for extracting elevation, vegetation and land surface temperature data 

In this tutorial, Connecting to the AppEEARS API, querying the list of available products, submitting an area sample request, downloading the request, working with the AppEEARS Quality API, and loading the results into R for visualization are covered. AppEEARS area sample requests allow users to subset their desired data by spatial area via vector polygons (shapefiles or GeoJSONs). Users can also reproject and reformat the output data. AppEEARS returns the valid data from the parameters defined within the sample request.       

### Data Used in the Example:

- Data layers: 
    - NASA MEaSUREs Shuttle Radar Topography Mission (SRTM) Version 3 Digital Elevation Model  
        - [SRTMGL1_NC.003](https://doi.org/10.5067/MEaSUREs/SRTM/SRTMGL1.003), 30m, static: 'SRTM_DEM' 
    - Combined MODIS Leaf Area Index (LAI)  
        - [MCD15A3H.061](https://doi.org/10.5067/MODIS/MCD15A3H.061), 500m, 4 day: 'Lai_500m'      
    - Terra MODIS Land Surface Temperature    
        - [MOD11A2.061](https://doi.org/10.5067/MODIS/MOD11A2.061), 1000m, 8 day: 'LST_Day_1km', 'LST_Night_1km'  

## Topics Covered in this Tutorial

1. **Getting Started**             
  1a. Load Packages                       
  1b. Set Up the Output Directory              
  1c. Login              
2. **Query Available Products**             
  2a. Search and Explore Available Products           
  2b. Search and Explore Available Layers             
3. **Submit an Area Request**               
  3a. Load a Shapefile            
  3b. Search and Explore Available Projections           
  3c. Compile a JSON Object                    
  3d. Submit a Task Request          
  3e. Retrieve Task Status           
4. **Download a Request**                       
  4a. Explore Files in Request Output             
  4b. Download Files in a Request (Automation)             
5. **Explore AppEEARS Quality Service**         
  5a. List Quality Layers             
  5b. Show Quality Values                     
  5c. Decode Quality Values                   
6. **BONUS: Load Request Output and Visualize**                          
  6a. Load a GeoTIFF                  
  6b. Plot a GeoTIFF                     

## Prerequisites and evironment Setup

This tutorial requires several R packages to be installed. The code below will check for the packages and install them if they are not already.

```{r}
# Required packages
packages <- c('earthdatalogin', 'getPass','httr','jsonlite','ggplot2','dplyr','tidyr','readr','geojsonio','geojsonR', 'sp', 'terra', 'rasterVis', 'RColorBrewer', 'jsonlite')

# Identify missing (not installed) packages
new.packages = packages[!(packages %in% installed.packages()[, "Package"])]

# Install new (not installed) packages
if(length(new.packages)) install.packages(new.packages, repos='http://cran.rstudio.com/', dependencies = TRUE) else print('All required packages are installed.')
```

***

## 1. Getting Started        

### 1a. Load Packages

First, load the R packages necessary to run the tutorial.

```{r, warning=FALSE, message=FALSE}
invisible(lapply(packages, library, character.only = TRUE))
```

### 1b. Set Up the Output Directory

Create an output directory for the results.

```{r, warning=FALSE, message=FALSE}
outDir <- file.path("../Data", "R_Output", fsep = "/")
dir.create(outDir)
```

### 1c. Login

Submitting a request requires you to authenticate using your NASA Earthdata Login username and password.

Start by assigning the AppEEARS API URL to a static variable. 

```{r}
API_URL = 'https://appeears.earthdatacloud.nasa.gov/api/'
```

Use the `httr` package to submit an HTTP POST request to login using the username and password. A successful login will provide you with a token to be used later in this tutorial to submit a request. For more information or if you are experiencing difficulties, please see the [API Documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#login).

```{r}
secret <- jsonlite::base64_enc(paste(
  getPass::getPass(msg = "Enter your NASA Earthdata Login Username:"),
  getPass::getPass(msg = "Enter your NASA Earthdata Login Password:"),
  sep = ":"))
```

Submit a POST request to the AppEEARS API's `login` service.

```{r}
login_req <- httr::POST(
  paste0(API_URL,"login"),
  httr:: add_headers(
    "Authorization" = paste("Basic", gsub("\n", "", secret)),
    "Content-Type" = "application/x-www-form-urlencoded;charset=UTF-8"),
  body = "grant_type=client_credentials")

httr::status_code(login_req)
```
A successful request will return `200` as the status.

Following a successful request, extract the content from the response and convern it to a JSON object to explore.

```{r}
# Retrieve the content of the request and convert the response to the JSON object
token_response <- toJSON(content(login_req), auto_unbox = TRUE)

# Print the prettified response
prettify(token_response)
```

Above, you should see a Bearer token. **This token is required to when interacting with the AppEEARS API.** It will expire approximately 48 hours after being acquired.

Let's save the token to a variable for use later.

```{r}
token <- paste("Bearer", fromJSON(token_response)$token)
```

## 2. Query Available Products

The product API provides details about all of the products and layers available in AppEEARS. For more information, please see the [API Documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#product).

Below, make a request to the `product` service to list all of the products available in AppEEARS.

```{r}
prod_req <- GET(paste0(API_URL, "product"))

# Retrieve the content of request and convert the info to JSON object
all_prods <- toJSON(content(prod_req), auto_unbox = TRUE)

# Print the prettified product response
prettify(all_prods)
```

### 2a. Search and Explore Available Products

Create a list indexed by product name to make it easier to query a specific product.

```{r}
# Divides information from each product.
divided_products <- split(fromJSON(all_prods), seq(nrow(fromJSON(all_prods))))

# Create a list indexed by the product name and version
products <- setNames(divided_products, fromJSON(all_prods)$ProductAndVersion)

# Print no. products available in AppEEARS
sprintf("AppEEARS currently supports %i products.", length(products))
```

Next, create a loop to get the product names and descriptions. Below, the 'ProductAndVersion' and 'Description' are printed for all products.

```{r}
for (p in products){
  print(paste0(p$ProductAndVersion," is ",p$Description," from ",p$Source))
}
```

The `product` service provides many useful details, including if a product is currently available in AppEEARS, a description, and information on the spatial and temporal resolution. Below, the product details are retrieved using 'ProductAndVersion'.

```{r}
# Convert the MCD15A3H.061 info to JSON object and print the prettified info
prettify(toJSON(products$"MCD15A3H.061"))
```

Also, the products can be searched using their description. Below, search for products containing Leaf Area Index (LAI) in their description and make a list of their ProductAndVersion.

```{r}
# Create an empty list
LAI_Products <- list()

# Loop through the product list and save all LAI products to LAI_Products
for (p in products){
  if (grepl('Leaf Area Index', p$Description )){
    LAI_Products <- append(LAI_Products, p$ProductAndVersion)
  }
}

LAI_Products
```

Using the info above, Create a list of desired products.  

```{r}
desired_products <- c('MCD15A3H.061','MOD11A2.061','SRTMGL1_NC.003')
desired_products
```

### 2b. Search and Explore Available Layers

This API call will list all of the layers available for a given product. Each product is referenced by its `ProductAndVersion` property which is also referred to as the product_id. First, request the layers for the `MCD15A3H.061` product.


```{r}
# Request layers for the 1st product in the list, i.e. MCD15A3H.061, from the product service
layer_req <- GET(paste0(API_URL,"product/", desired_products[1]))

# Retrieve content of the request and convert the content to JSON object
MCD15A3H_response <- toJSON(content(layer_req), auto_unbox = TRUE)
```

The response will contain the layer names and attributes for the product.

```{r}
# Print the prettified response
prettify(MCD15A3H_response)
```

To submit a request, the layer names are needed. Below, extract the layer names from the response.

```{r}
names(fromJSON(MCD15A3H_response))
```

Do the same for the `MOD11A2.061` product. Request the layers information and print the layer names.

```{r}
layer_req <- GET(paste0(API_URL,"product/", desired_products[2]))

# Retrieve content of the request and convert the content to JSON object
MOD11_response <- toJSON(content(layer_req), auto_unbox = TRUE)

# Print the layer names
names(fromJSON(MOD11_response))
```

Finally, request the layers for the`SRTMGL1_NC.003` product.

```{r}
layer_req <- GET(paste0(API_URL,"product/", desired_products[3]))

# Retrieve content of the request and convert the content to JSON object
SRTMGL1_response <- toJSON(content(layer_req), auto_unbox = TRUE)

# Print the layer names
names(fromJSON(SRTMGL1_response))
```

Now we will select the desired layers and pertinent products and make a dataframe using this information. This list will be inserted into the JSON file used to submit a request in Section 3.

```{r}
# Create a vector of desired layers
desired_layers <- c("LST_Day_1km", "LST_Night_1km", "Lai_500m", "SRTMGL1_DEM")

# Create a vector of products including the desired layers
desired_prods <- c("MOD11A2.061", "MOD11A2.061", "MCD15A3H.061", "SRTMGL1_NC.003") 

# Create a dataframe including the desired data products and layers
layers <- data.frame(product = desired_prods, layer = desired_layers)
```

## 3. Submit an Area Request

The Submit task API call provides a way to submit a new request to be processed. It can accept data via JSON or query string. In the example below, create a JSON object and submit a request. Tasks in AppEEARS correspond to each request associated with your user account. Therefore, each of the calls to this service requires an authentication token (see Section 1c.).

### 3a. Load a Shapefile

In this section, begin by loading a shapefile using the `rgdal` package. The shapefile is publicly available for download from the [NPS website](https://irma.nps.gov/DataStore/Reference/Profile/2224545?lnv=True).

```{r}
data_path = 'Data'
shapefile_loc = paste0(data_path, '/nps_boundary.shp')
nps <- terra::vect(shapefile_loc)
head(nps)
```

Below, subset the shapefile for the national park that you are interested in using for your region of interest, here Grand Canyon National Park.

```{r}
# subset the shapefile to keep Grand Canyon
nps_gc <- nps[nps$UNIT_NAME == "Grand Canyon National Park",]
remove(nps)
```

**nps_gc** is a `SpatVector` class from terra. We need to convert this to an `sf` object and then to a GeoJSON object.

```{r, message= FALSE}
nps_gc_json <- geojsonio::geojson_json(sf::st_as_sf(nps_gc))
remove(nps_gc)
```

Next, the GeoJSON object can be read as a list using the `FROM_GeoJson` function from `geojsonR` package. 

```{r}
nps_gc_js <- geojsonR::FROM_GeoJson(nps_gc_json)
remove(nps_gc_json)
```

### 3b. Search and Explore Available Projections

The spatial API provides some helper services used to support submitting area task requests. The call below will retrieve the list of supported projections in AppEEARS. For more information, please see the [AppEEARS API](https://appeears.earthdatacloud.nasa.gov/api/?language=R#spatial)

```{r}
# Request the projection info from API_URL
proj_req <- GET(paste0(API_URL, "spatial/proj"))

# Retrieve content of the request and convert to JSON object
proj_response <- toJSON(content(proj_req), auto_unbox = TRUE)

# List the available projections
projs <- fromJSON(proj_response)
projs
```

In this example we will use the geographic projection. Below, subset the projection dataframe to select the geographic projection.

```{r}
projection <- projs[projs$Name=="geographic",]
```

### 3c. Compile a JSON Object

In this section, begin by setting up the information needed for a nested dataframe that will be later converted to a JSON object for submitting an AppEEARS area request. For detailed information on required JSON parameters, see the [API Documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#tasks).

We'll start by specifying the task name, task type (e.g., point or area), output projection, output format (e.g., geotiff or netcdf4), and date range for the request.

```{r}
taskName <- 'Grand Canyon'
taskType <- 'area'
projection <- projection$Name
outFormat <- 'geotiff'
# Start of the date range for which to extract data: MM-DD-YYYY
startDate <- '01-01-2018'
# End of the date range for which to extract data: MM-DD-YYYY
endDate <- '12-31-2019'
```

If you are interested in submitting a request using a recurring time in a year, set the time variables as below:

```{r}
# startDate <- '05-01'                 # change start/end date to MM-DD
# endDate <- '06-30'
# recurring <- TRUE                    # Specify True for a recurring date range
# fromYear <- 2018
# toYear <- 2020
```

To submit a task successfully, a specifically formatted JSON object is required. The following code creates a nested dataframe using previous information, which will be converted into the necessary JSON object. 

```{r}
# Create a dataframe including the date range for the request.
date <- data.frame(startDate = startDate, endDate = endDate)

# If you set the recurring to TRUE
#date <- data.frame(startDate = startDate, endDate = endDate , recurring = recurring)
#date$yearRange <- list(c(fromYear,toYear))
```

Next, create a list including the projection and add the output format information.

```{r}
out <- list(projection)
names(out) <- c("projection")
out$format$type <- outFormat
```

Change the GeoJSON format for successful task submission.

```{r}
nps_gc_js$features[[1]]$geometry$coordinates <- list(nps_gc_js$features[[1]]$geometry$coordinates)
```

Next, compile dataframes and lists to create a nested dataframe. 

```{r}
# Create a list of dataframes and assign names
task_info <- list(date, layers, out, nps_gc_js)
names(task_info) <- c("dates", "layers", "output", "geo")

# Create the nested list and assing names
task <- list(task_info, taskName, taskType)
names(task) <- c("params", "task_name", "task_type") 
```

`toJSON` function from `jsonlite` package converts the type of dataframe to a string that can be recognized as a JSON object to be submitted as an area request.

```{r}
task_json <- jsonlite::toJSON(task, auto_unbox = TRUE, digits = 10)
```

### 3d. Submit a Task Request 

Now, let's send an HTTP POST request to the AppEEARS API task service using the token we previously saved to authenticate the request. The `task_json` is passed as the body of the request. 

```{r}
# Post the point request to the API task service
response <- POST(
  paste0(API_URL, "task"),
  body = task_json,
  encode = "json",
  httr::add_headers(Authorization = token, "Content-Type" = "application/json"))

httr::status_code(response)
```

A status code of **202** should be returned.

```{r}
# Retrieve content of the request
task_response <- jsonlite::toJSON(content(response), auto_unbox = TRUE)

# Print the response
prettify(task_response)
```

All AppEEARS requests generate a JSON output containing the input parameters submitted for a request. The output JSON file can be read in and used to submit request using the code below. This is useful if someone share an AppEEARS JSON file for you to reproduce the results.

```{r}
# task <- jsonlite::toJSON(jsonlite::read_json("LST-request.json"), digits = 10, auto_unbox = TRUE)
# response <- POST(paste0(API_URL,"task"), body = task, encode = "json", 
#                        add_headers(Authorization = token, "Content-Type" = "application/json"))
# task_response <- prettify(toJSON(content(response), auto_unbox = TRUE))
# task_response
```

### 3e. Retrieve Task Status

This API call will list all of the requests associated with your user account, automatically sorted by date descending with the most recent requests listed first.
The AppEEARS API contains some helpful formatting resources. Below, limit the API response to 2 entries for the last 2 requests and set pretty to True to format the response as an organized JSON, making it easier to read. Additional information on AppEEARS API
[retrieve task](https://appeears.earthdatacloud.nasa.gov/api/?language=R#retrieve-task), [pagination](https://appeears.earthdatacloud.nasa.gov/api/?language=R#pagination), and [formatting](https://appeears.earthdatacloud.nasa.gov/api/?language=R#formatting) can be found in the API documentation.

```{r}
params <- list(limit = 2, pretty = TRUE)
# Request the task status of last 2 requests from task URL
response_req <- GET(
  paste0(API_URL, "task"),
  query = params,
  httr::add_headers(Authorization = token))

# Retrieve content of the request as JSON
status_response <- toJSON(content(response_req), auto_unbox = TRUE)

# Print the prettified response
prettify(status_response)
```

A **task id** is generated when submitting your request. The **task id** can be used to check the status of the request and is also used when indentifing and accessing results from the request. Below, the **task id** is extracted from the response in this example.

> **NOTE**: The **task id** from any request can be substituded in the code below to get information (e.g., status or bundle content) about the the request.

```{r}
task_id <- fromJSON(task_response)[[1]]

```

# Retrieve the status of the request using the **task_id**.

```{r}
## Request the task status of a task with the provided task_id from task URL
#status_req <- GET(paste0(API_URL,"task/", task_id), add_headers(Authorization = token))

## Retrieve content of the request
#statusResponse <-toJSON(content(status_req), auto_unbox = TRUE)

## Print the prettified response
#prettify(statusResponse)
```

Some request will take a bit of time to complete. We can set up a loop to retrieve the task status every 60 seconds. When the task status is `done` the content for our request will be available to access or download.

```{r}
# Request the task status of last request from task URL
stat_req <- GET(
  paste0(API_URL,"task"),
  query = list(limit = 1),
  httr::add_headers(Authorization = token))

# Retrieve content of the request as JSON
stat_response <- toJSON(content(stat_req), auto_unbox = TRUE)

# Assign the task status to a variable
stat <- fromJSON(stat_response)$status

while (stat != 'done') {
  Sys.sleep(60)
  stat_req <- GET(
    paste0(API_URL,"task"),
    query = list(limit = 1),
    httr::add_headers(Authorization = token))
  stat <- fromJSON(toJSON(content(stat_req), auto_unbox = TRUE))$status
  print(stat)
}
```

## 4. Download a Request

### 4a. Explore Files in Request Output

Before downloading the request output, examine the files contained in the request output.

```{r}
task_id <- fromJSON(task_response)[[1]]
response <- GET(paste0(API_URL, "bundle/", task_id), add_headers(Authorization = token))
bundle_response <- prettify(toJSON(content(response), auto_unbox = TRUE))
```

### 4b. Download Files in a Request (Automation) 

The bundle API provides information about completed tasks. For any completed task, a bundle can be queried to return the files contained as a part of the task request. Below, call the bundle API and return all of the output files. Next, read the contents of the bundle in JSON format and loop through file_id to automate downloading all of the output files into the output directory. For more information, please see [AppEEARS API Documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#bundle).

```{r, results= 'hide'}
bundle <- fromJSON(bundle_response)$files
for (id in bundle$file_id){
  # Retrieve the filename from the file_id
  filename <- bundle[bundle$file_id == id,]$file_name
  # Create a destination directory to store the file in
  filepath <- paste(outDir,filename, sep = "/")
  suppressWarnings(dir.create(dirname(filepath)))
  # Write the file to disk using the destination directory and file name
  response <- GET(paste0(API_URL, "bundle/", task_id, "/", id), 
                  write_disk(filepath, overwrite = TRUE), progress(),
                  add_headers(Authorization = token))
}
```

Now, take a look at "R_output" directory. Separate folders are made for each product and the outputs are saved in these folders. 

```{r}
# List of directories in the R_output directory
list.dirs(outDir)
```

Below, the list of relative path and file names is assigned to a variable and part of the list is printed.

```{r}
# Assign relative path to a variable
relative_path <- bundle$file_name
# Print part of the list
relative_path[550:560]
```

Later in this tutorial, the `SRTMGL1_NC` GeoTIFF is loaded for visulalization. Below, the directory to this file is assigned to a variable.

```{r}
# Extract the telative path to the SRTMGL1_N
SRTMGL1_NC_subdir <- relative_path[grepl("*SRTMGL1_NC", relative_path)]
# Assign absolute path to a variable
SRTMGL1_NC_dir <- paste0(outDir, "/", SRTMGL1_NC_subdir)
# Print the absolute path
SRTMGL1_NC_dir
```

## 5. Explore AppEEARS Quality Service

The quality API provides quality details about all of the data products available in AppEEARS. Below are examples of how to query the quality API for listing quality products, layers, and values. The final example (Section 5c.) demonstrates how AppEEARS quality services can be leveraged to decode pertinent quality values for your data. For more information visit [AppEEARS API documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#quality)

First, reset pagination to include offset which allows you to set the number of results to skip before starting to return entries. Next, make a call to list all of the data product layers and the associated quality product and layer information.

```{r}
# Assign query to a variable
params <- list(limit = 6, offset = 20, pretty = TRUE)
# Request the specified quality layer info from quality API
quality_req <- GET(paste0(API_URL, "quality"), query = params)
# Retrieve the content of request
quality_response <- toJSON(content(quality_req), auto_unbox = TRUE)
# Remove the variables that are not needed and print the quality information
remove(quality_req, quality_content)
prettify(quality_response)
```

### 5a. List Quality Layers

This API call will list all of the quality layer information for a product. For more information visit [AppEEARS API documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#quality)

```{r}
# Assign productAndVersion to a variable
productAndVersion <- 'MCD15A3H.006'
# Request quality info for a product from quality API
MCD15A3H_q_req <- GET(paste0(API_URL, "quality/", productAndVersion))
# Retrieve the content of request
MCD15A3H_quality <- toJSON(content(MCD15A3H_q_req), auto_unbox = TRUE)
# Remove the variables that are not needed and print the quality information
remove(MCD15A3H_q_req, MCD15A3H_q_content)
prettify(MCD15A3H_quality)
```

### 5b. Show Quality Values 

This API call will list all of the values for a given quality layer. 

```{r}
# Assign a quality layer to a variable
quality_layer <- 'FparLai_QC'
# Request the specified quality layer info from quality API
quality_req <- GET(paste0(API_URL, "quality/",  productAndVersion, "/", quality_layer, sep = ""))
# Retrieve the content of request
quality_response <- toJSON(content(quality_req), auto_unbox = TRUE)
# Remove the variables that are not needed and print the quality response
remove(quality_req, quality_content)
prettify(quality_response)
```

### 5c. Decode Quality Values

This API call will decode the bits for a given quality value. 

```{r}
# Assign a quality value to a variable
quality_value <- 1
# Request and retrieve information for provided quality value from quality API URL 
response <- content(GET(paste0(API_URL, "quality/", productAndVersion, "/", quality_layer, "/", quality_value)))
# Convert the info to JSON object
q_response <- toJSON(response, auto_unbox = TRUE)
# Remove the variables that are not needed anymore and print the quality response
remove(response)
prettify(q_response)
```

## 6. **BONUS: Load Request Output and Visualize**

Here, load one of the output GeoTIFFs and show some basic visualizations using the `rasterVis` and `ggplot2` packages.

### 6a. Load a GeoTIFF

First, created a raster object by calling the raster() function from the `raster` package.

```{r}
# Create a raster object using the directory previously extracted
dem <- terra::rast(SRTMGL1_NC_dir)
```

### 6b. Plot a GeoTIFF

Make a plot of DEM data and add some additional parameters to the plot. 

```{r, warning= FALSE, fig.width = 8, fig.height=5}
gplot(dem) +
  geom_raster(aes(fill = value)) +
  scale_fill_distiller(name = "Elevation (m)", palette = "BrBG", na.value=NA) +
  coord_fixed(expand = F, ratio = 1) +
  labs(title = "SRTM DEM: Grand Canyon NP")+
  theme(plot.title = element_text(face = "bold",size = rel(2),hjust = 0.5),
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        panel.border = element_rect(fill = NA , color = "black", size = 0.8),
        panel.background = element_rect(fill = "white", colour = "#6D9EC1",size = 2, linetype = "solid"),
        panel.grid.major = element_line(size = 0.001, linetype = 'solid',colour = "gray"),
        panel.grid.minor = element_line(size = 0.001, linetype = 'solid',colour = "gray"))
```

This example can provide a template to use for your own research workflows. Leveraging the AppEEARS API for searching, extracting, and formatting analysis ready data, and loading it directly into R means that you can keep your entire research workflow in a single software program, from start to finish.

***

## Contact Information  
 Material written by LP DAAC^1^                                                     
 Contact: LPDAAC@usgs.gov   
 Voice: +1-866-573-3222   
 Organization: Land Processes Distributed Active Archive Center (LP DAAC)  
 Website: https://lpdaac.usgs.gov/  
 Date last modified: 11-04-2024      
 
 ^1^ KBR, Inc., contractor to the U.S. Geological Survey, Earth Resources Observation and Science (EROS) Center,  
 Sioux Falls, South Dakota, USA. Work performed under USGS contract G15PD00467 for LP DAAC^2^.  
 
 ^2^ LP DAAC Work performed under NASA contract NNG14HH33I.
