---
title: "Getting Started with the AppEEARS API (Point Request)"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    theme: paper
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
knitr::opts_knit$set(root.dir = dirname(rprojroot::find_rstudio_root_file()))
```

***

**This tutorial demonstrates how to use R to connect to the AppEEARS API**

The Application for Extracting and Exploring Analysis Ready Samples ([AppEEARS](https://appeears.earthdatacloud.nasa.gov/)) offers a simple and efficient way to access and transform [geospatial data](https://appeears.earthdatacloud.nasa.gov/products) from a variety of federal data archives in an easy-to-use web application interface. AppEEARS enables users to subset geospatial data spatially, temporally, and by band/layer for point and area samples. AppEEARS returns not only the requested data, but also the associated quality values, and offers interactive visualizations with summary statistics in the web interface. The [AppEEARS API](https://appeears.earthdatacloud.nasa.gov/api/) offers users **programmatic access** to all features available in AppEEARS, with the exception of visualizations. The API features are demonstrated in this tutorial.

## Example: Submit a point request with multiple points in U.S. National Parks for extracting vegetation and land surface temperature data  

In this tutorial, Connecting to the AppEEARS API, querying the list of available products, submitting a point sample request, downloading the request, working with the AppEEARS Quality API, and loading the results into R for visualization are covered. AppEEARS point requests allow users to subset their desired data using latitude/longitude geographic coordinate pairs (points) for a time period of interest, and for specific data layers within data products. AppEEARS returns the valid data from the parameters defined within the sample request.

### Data Used in the Example:

- Data layers: 
    - Combined MODIS Leaf Area Index (LAI)  
        - [MCD15A3H.061](https://doi.org/10.5067/MODIS/MCD15A3H.061), 500m, 4 day: 'Lai_500m'      
    - Terra MODIS Land Surface Temperature    
        - [MOD11A2.061](https://doi.org/10.5067/MODIS/MOD11A2.061), 1000m, 8 day: 'LST_Day_1km', 'LST_Night_1km'

## Topics Covered in this tutorial:

1. **Getting Started**  
    1a. Load Packages
    1b. Set Up the Output Directory  
    1c. Login   
2. **Query Available Products**  
    2a. Search and Explore Available Products   
    2b. Search and Explore Available Layers     
3. **Submit a Point Request**  
    3a. Compile a JSON Object 
    3b. Submit a Task Request 
    3c. Retrieve Task Status  
4. **Download a Request**    
    4a. Explore Files in Request Output   
    4b. Download Files in a Request (Automation)  
5. **Explore AppEEARS Quality API**  
    5a. List Quality Layers    
    5b. Show Quality Values  
    5c. Decode Quality Values  
6. **BONUS: Load Request Output and Visualize**    
    6a. Load CSV     
    6b. Plot Results (Line/Scatter Plots)  

## Prerequisites and evironment Setup

This tutorial requires the following R packages to be installed. The code below will check for the packages and install them if they are not already.

```{r}
# Required packages
packages <- c('earthdatalogin', 'getPass','httr','jsonlite','ggplot2','dplyr','tidyr','readr','geojsonio','geojsonR', 'sp', 'terra', 'rasterVis', 'RColorBrewer', 'jsonlite')

# Identify missing (not installed) packages
new.packages = packages[!(packages %in% installed.packages()[, "Package"])]

# Install new (not installed) packages
if(length(new.packages)) install.packages(new.packages, repos='http://cran.rstudio.com/', dependencies = TRUE) else print('All required packages are installed.')
```

***

## 1. Getting Started

### 1a. Load Packages

First, load the R packages necessary to run the tutorial.

```{r, warning=FALSE, message=FALSE}
invisible(lapply(packages, library, character.only = TRUE))
```

### 1b. Set Up the Output Directory

Set your input directory, and create an output directory for the results.

```{r, warning=FALSE, message=FALSE}
outDir <- file.path("../Data", "R_Output", fsep="/")
dir.create(outDir)
```

### 1c. [Login]{#login}

Submitting a request requires you to authenticate using your NASA Earthdata Login username and password.

Start by assigning the AppEEARS API URL to a static variable. 

```{r}
API_URL = 'https://appeears.earthdatacloud.nasa.gov/api/'
```

Use the `httr` package to submit an HTTP POST request to login using the username and password. A successful login will provide you with a token to be used later in this tutorial to submit a request. For more information or if you are experiencing difficulties, please see the [API Documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#login).

```{r}
secret <- jsonlite::base64_enc(paste(
  getPass::getPass(msg = "Enter your NASA Earthdata Login Username:"),
  getPass::getPass(msg = "Enter your NASA Earthdata Login Password:"),
  sep = ":"))
```

 Submit a POST request to the AppEEARS API's `login` service.

```{r}
login_req <- httr::POST(
  paste0(API_URL,"login"),
  httr:: add_headers(
    "Authorization" = paste("Basic", gsub("\n", "", secret)),
    "Content-Type" = "application/x-www-form-urlencoded;charset=UTF-8"),
  body = "grant_type=client_credentials")

httr::status_code(login_req)
```
A successful request will return `200` as the status.

Following a successful request, extract the content from the response and convern it to a JSON object to explore.

```{r}
# Retrieve the content of the request and convert the response to the JSON object
token_response <- toJSON(content(login_req), auto_unbox = TRUE)

# Print the prettified response
prettify(token_response)
```

Above, you should see a Bearer token. **This token is required to when interacting with the AppEEARS API.** It will expire approximately 48 hours after being acquired.

## 2. Query Available Products

The product API provides details about all of the products and layers available in AppEEARS. For more information, please see the [API Documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#product).

Below, make a request to the `product` service to list all of the products available in AppEEARS.

```{r}
prod_req <- GET(paste0(API_URL, "product"))

# Retrieve the content of request and convert the info to JSON object
all_prods <- toJSON(content(prod_req), auto_unbox = TRUE)

# Print the prettified product response
prettify(all_prods)
```

### 2a. Search and Explore Available Products

Create a list indexed by product name to make it easier to query a specific product.

```{r}
# Divides information from each product.
divided_products <- split(fromJSON(all_prods), seq(nrow(fromJSON(all_prods))))

# Create a list indexed by the product name and version
products <- setNames(divided_products, fromJSON(all_prods)$ProductAndVersion)

# Print no. products available in AppEEARS
sprintf("AppEEARS currently supports %i products.", length(products))
```

Next, create a loop to get the product names and descriptions. Below, the 'ProductAndVersion' and 'Description' are printed for all products.

```{r}
for (p in products){
  print(paste0(p$ProductAndVersion, " is ", p$Description, " from ", p$Source))
}
```

The `product` service provides many useful details, including if a product is currently available in AppEEARS, a description, and information on the spatial and temporal resolution. Below, the product details are retrieved using 'ProductAndVersion'.

```{r}
# Convert the MCD15A3H.061 info to JSON object and print the prettified info
prettify(toJSON(products$"MCD15A3H.061"))
```

Also, the products can be searched using their description. Below, search for products containing Leaf Area Index (LAI) in their description and make a list of their ProductAndVersion.

```{r}
# Create an empty list
LAI_Products <- list()

# Loop through the product list and save all LAI products to LAI_Products
for (p in products){
  if (grepl('Leaf Area Index', p$Description )){
    LAI_Products <- append(LAI_Products, p$ProductAndVersion)
  }
}

LAI_Products
```

Using the info above, create a list of desired products.  

```{r}
desired_products <- c('MCD15A3H.061','MOD11A2.061')
desired_products
```

### 2b. Search and Explore Available Layers 

This API call will list all of the layers available for a given product. Each product is referenced by its `ProductAndVersion` property which is also referred to as the product_id. First, request the layers for the `MCD15A3H.061` product.

```{r}
# Request layers for the 1st product in the list, i.e. MCD15A3H.061, from the product service
layer_req <- GET(paste0(API_URL,"product/", desired_products[1]))

# Retrieve content of the request and convert the content to JSON object
MCD15A3H_response <- toJSON(content(layer_req), auto_unbox = TRUE)
```

The response will contain the layer names and attributes for the product.

```{r}
# Print the prettified response
prettify(MCD15A3H_response)
```

To submit a request, the layer names are needed. Below, extract the layer names from the response.

```{r}
names(fromJSON(MCD15A3H_response))
```

Do the same for the `MOD11A2.061` product. Request the layers information and print the layer names.

```{r}
layer_req <- GET(paste0(API_URL,"product/", desired_products[2]))

# Retrieve content of the request and convert the content to JSON object
MOD11_response <- toJSON(content(layer_req), auto_unbox = TRUE)

# Print the layer names
names(fromJSON(MOD11_response))
```

Finally, select the desired layers and pertinent products and make a dataframe using this information. This dataframe will be used to construct a JSON object to submit a request in [Section 3](#section3).

```{r}
# Create a vector of products including the desired layers
desired_prods <- c("MOD11A2.061", "MOD11A2.061", "MCD15A3H.061")

# Create a vector of desired layers
desired_layers <- c("LST_Day_1km", "LST_Night_1km", "Lai_500m")

# Create a dataframe including the desired data products and layers
layers <- data.frame(product = desired_prods, layer = desired_layers)
```

## 3. [Submit a Point Request]{#section3}

The Submit task API call provides a way to submit a new request to be processed. It can accept data via JSON or query string. In the example below, create a JSON object and submit a request. Tasks in AppEEARS correspond to each request associated with your user account. Therefore, each of the calls to this service requires an authentication token (see [Section 1c.](#login)).

### 3a. Compile a JSON Object

In this section, begin by setting up the information needed for a nested dataframe that will be later converted to a JSON object for submitting an AppEEARS point request. For detailed information on required JSON parameters, see the [API Documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#tasks).

For point requests, beside the date range and desired layers information, the coordinates property must also be inside the task object. Optionally, set `id` and `category` properties to further identify your selected coordinates.

```{r}
# Start of the date range for which to extract data: MM-DD-YYYY
startDate <- "01-01-2018"
# End of the date range for which to extract data: MM-DD-YYYY
endDate <- "12-31-2018"

# If you are interested in submitting a request using a recurring time in a year, set the time variables as below:
# startDate <- '05-01'                 # change start/end date to MM-DD
# endDate <- '06-30'
# recurring <- TRUE                    # Specify True for a recurring date range
# fromYear <- 2018
# toYear <- 2020

lat <- c(36.206228, 37.289327)        # Latitude of the point sites
lon <- c(-112.127134, -112.973760)    # Longitude of the point sites

id <- c("0","1")                      # ID for the point sites
category <- c("Grand Canyon", "Zion") # Category for point sites

taskName <- 'NPS Vegetation'           # Enter name of the task, 'NPS Vegetation' used here
taskType <- 'point'                    # Specify the task type, it can be either "area" or "point"
```

To be able to successfully submit a task, the JSON object should be structured in a certain way. The code chunk below uses the information from the previous chunk to create a nested dataframe. This nested dataframe will be converted to JSON object that can be used to complete the request. 

```{r}
# Create a dataframe including the date range for the request
date <- data.frame(startDate = startDate, endDate = endDate)

# If you set the recurring to TRUE
# date <- data.frame(startDate = startDate, endDate = endDate , recurring = recurring)
# date$yearRange <- list(c(fromYear,toYear))

# Create a dataframe including lat and long coordinates. ID and category name is optional.
coordinates <- data.frame(id = id, longitude = lon, latitude = lat, category = category)

# Create a list of dataframes and assign names
task_info <- list(date, layers, coordinates)
names(task_info) <- c("dates", "layers", "coordinates")

# Create a final list of dataframes and assign names
task <- list(task_info, taskName, taskType)
names(task) <- c("params", "task_name", "task_type")
```

`toJSON` function from `jsonlite` package converts the type of dataframe to a string that can be recognized as a JSON object to be submitted as a point request.

```{r}
task_json <- toJSON(task,auto_unbox = TRUE)
prettify(task_json)
```

### 3b. Submit a Task Request
Token information is needed to submit a request. Below the login token is assigned to a variable.

```{r}
token <- paste("Bearer", fromJSON(token_response)$token)     # Save login token to a variable
```

Below, submit a POST request to the API **task service**, using the `task_json` created above.

```{r}
# Post the point request to the API task service
response <- httr::POST(
  paste0(API_URL, "task"),
  body = task_json,
  encode = "json",
  httr::add_headers(Authorization = token, "Content-Type" = "application/json"))

# Print the status code of the response
httr::status_code(response)
```

```{r}
# Retrieve content of the request and convert the content to JSON object
task_response <- prettify(toJSON(content(response), auto_unbox = TRUE))

# Print the prettified task response
task_response
```

### 3c. Retrieve Task Status

This API call will list all of the requests associated with your user account, automatically sorted by date descending with the most recent requests listed first.
The AppEEARS API contains some helpful formatting resources. Below, limit the API response to 2 entries for the last 2 requests and set pretty to True to format the response as an organized JSON object to make it easier to read. Additional information on AppEEARS API
[retrieve task](https://appeears.earthdatacloud.nasa.gov/api/?language=R#retrieve-task), [pagination](https://appeears.earthdatacloud.nasa.gov/api/?language=R#pagination), and [formatting](https://appeears.earthdatacloud.nasa.gov/api/?language=R#formatting) can be found in the API documentation.

```{r}
# Request the task status of last 2 requests from task URL
response_req <- GET(
  paste0(API_URL, "task"),
  add_headers(Authorization = token))

# Retrieve content of the request
response_content <- content(response_req)
```

The response above contains information for every request associated with your user account. Below, look at the first two requests, which should be the most recent two requests.  

```{r}
str(response_content[1:2], max.level = 3)
```

The task_id is needed to retrieve information or content for a specific request. Previously, the task_id was generated when submitting your request. This can be used to retrieve a status about the request.
 
```{r}
# Extract the task_id of submitted point request
task_id <- fromJSON(task_response)$task_id
# Request the task status of a task with the provided task_id from task URL
status_req <- GET(
  paste0(API_URL, "task/", task_id),
  add_headers(Authorization = token))

# Retrieve content of the request and convert the content to JSON object
status_response <-toJSON(content(status_req), auto_unbox = TRUE)

# Get the status of the task
status <- fromJSON(status_response)$status

# Print the status
print(status)
```

Retrieve the task status every 5 seconds. The task status should be `done` to be able to download the output.

```{r}
while (status != 'done') {
  Sys.sleep(5)
  # Request the task status and retrieve content of request from task URL
  stat_content <- content(GET(
    paste0(API_URL,"task/", task_id),
    add_headers(Authorization = token)))
  
  status <-fromJSON(toJSON(stat_content, auto_unbox = TRUE))$status    # Get the status
  remove(stat_content)
  print(status)
}
```

## 4. Download a Request

### 4a. Explore Files in Request Output 

Before downloading the request output, examine the files contained in the request output.

```{r}
# Request the task bundle info from API bundle URL
response <- GET(
  paste0(API_URL, "bundle/", task_id),
  add_headers(Authorization = token))

# Retrieve content of the request convert the content to JSON object
bundle_response <- toJSON(content(response), auto_unbox = TRUE)

# Print the prettified bundle response
prettify(bundle_response)
```

### 4b. Download Files in a Request (Automation)

The bundle API provides information about completed tasks. For any completed task, a bundle can be queried to return the files contained as a part of the task request. Below, call the bundle API and return all of the output files. Next, read the contents of the bundle in JSON format and loop through file_id to automate downloading all of the output files into the output directory. For more information, please see [AppEEARS API Documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#bundle).

```{r, warning= FALSE, results= 'hide'}
bundle <- fromJSON(bundle_response)$files

for (id in bundle$file_id){
  # Retrieve the filename from the file_id
  filename <- bundle[bundle$file_id == id,]$file_name

  # Create a destination directory to store the file in
  filepath <- paste(outDir, filename, sep = "/")
  suppressWarnings(dir.create(dirname(filepath)))

  # Write the file to disk using the destination directory and file name
  response <- GET(
    paste0(API_URL, "bundle/", task_id, "/", id),
    write_disk(filepath, overwrite = TRUE),
    progress(),
    add_headers(Authorization = token))
}
```

## 5. Explore AppEEARS Quality Service

The quality API provides quality details about all of the data products available in AppEEARS. Below are examples of how to query the quality API for listing quality products, layers, and values. The final example ([Section 5c.](#decode)) demonstrates how AppEEARS quality services can be leveraged to decode pertinent quality values for your data. For more information visit [AppEEARS API documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#quality).

First, reset pagination to include offset which allows you to set the number of results to skip before starting to return entries. Next, make a call to list all of the data product layers and the associated quality product and layer information.

```{r}
# Request the quality info from quality API_URL
q_req <- GET(paste0(API_URL, "quality"))

# Retrieve the content of request and convert the content to JSON object
q_response <- toJSON(content(q_req), auto_unbox = TRUE)

# Print the prettified quality information
prettify(q_response)
```

### 5a. List Quality Layers

This API call will list all of the quality layer information for a product. For more information visit [AppEEARS API documentation](https://appeears.earthdatacloud.nasa.gov/api/?language=R#quality)

```{r}
# Assign ProductAndVersion to a variable
productAndVersion <- 'MCD15A3H.061'

# Request the quality info from quality API for a specific product
MCD15A3H_req <- GET(paste0(API_URL, "quality/", productAndVersion))

# Retrieve the content of request and convert the content to JSON object
MCD15A3H_quality <- toJSON(content(MCD15A3H_req), auto_unbox = TRUE)

# Print the quality information
prettify(MCD15A3H_quality)
```

### 5b. Show Quality Values

This API call will list all of the values for a given quality layer.

```{r}
# Assign a quality layer to a variable
quality_layer <- 'FparLai_QC'

# Request the specified quality layer info from quality API
quality_req <- GET(paste0(API_URL, "quality/",  productAndVersion, "/", quality_layer, sep = ""))

# Retrieve the content of request and convert the content to JSON object
quality_response <- toJSON(content(quality_req), auto_unbox = TRUE)

# Print the quality response as a dataframe
prettify(quality_response)
```

### 5c. [Decode Quality Values]{#decode}

This API call will decode the bits for a given quality value. 

```{r}
quality_value <- 1

# Request and retrieve information for provided quality value from quality API URL 
response <- content(GET(paste0(API_URL, "quality/", productAndVersion, "/", quality_layer, "/", quality_value)))

# Convert the content to JSON object
q_response <- toJSON(response, auto_unbox = TRUE)

# Print the response
prettify(q_response)
```

## 6. **BONUS**: Load Request Output and Visualize

Here, load the CSV file containing the results from your request using `readr` package, and create some basic visualizations using the `ggplot2` package.

### 6a. Load a CSV

Use the `readr` package to load the CSV file containing the results from the AppEEARS request.

```{r, message= FALSE, warning= FALSE}
# Make a list of csv files in the output directory
files <- list.files(outDir, pattern = "\\MOD11A2-061-results.csv$")

# Read the MOD11A2 results
df <- read_csv(paste0(outDir,"/", files), show_col_types = FALSE)
```

Select the MOD11A2.061 LST Day column for the data from Grand Canyon National Park using `dplyr` package.

```{r}
lst_GC <- df %>%
  # Filter df for the point from GC
  filter(Category == "Grand Canyon") %>%  
  # Select desired columns
  select(Latitude, Longitude, Date ,MOD11A2_061_LST_Day_1km, MOD11A2_061_LST_Night_1km)
```

Extract information for LST_DAY_1KM from MOD11_response of product service call from earlier in the tutorial.

```{r}
# Extract all of the data layer info for LST_Day_1km
#fromJSON(MOD11_response)$LST_Day_1km

# Assign fill value and units to a variable
fill_value <- fromJSON(MOD11_response)$LST_Day_1km$FillValue
unit <- fromJSON(MOD11_response)$LST_Day_1km$Units

# Print the fill value and unit
sprintf("Fill value for LST_DAY_1KM is: %i", fill_value)
sprintf("Unit for LST_DAY_1KM is: %s", unit)

```

### 6b. Plot Results (Line/Scatter Plots)

Next, plot a time series of daytime LST for the selected point in Grand Canyon National Park for 2018. Below, filter the LST data to exclude fill values. 

```{r}
lst_GC <- lst_GC %>%
  # exclude NoData
  filter(MOD11A2_061_LST_Day_1km != fill_value) %>%
  filter(MOD11A2_061_LST_Night_1km != fill_value)

```

Next, plot LST Day as a time series with some additional formatting using `ggplot2`.

```{r, fig.width = 12, fig.height=5}
ggplot(lst_GC)+
  geom_line(aes(x= Date, y = MOD11A2_061_LST_Day_1km), size=1, color="blue")+
  geom_point(aes(x= Date, y = MOD11A2_061_LST_Day_1km), shape=18 , size = 3, color="blue")+
  labs(title = "Time Series",
       x = "Date",
       y = sprintf( "LST_Day_1km (%s)", unit))+
  scale_x_date(date_breaks = "16 day", limits = as.Date(c('2018-01-01','2019-01-01')))+ 
  scale_y_continuous(limits = c(250, 325), breaks = seq(250, 325, 10))+
  theme(plot.title = element_text(face = "bold",size = rel(2.5),hjust = 0.5),
        axis.title = element_text(face = "bold",size = rel(1)),
        panel.background = element_rect(fill = "lightgray", colour = "black"),
        axis.text.x = element_text(face ="bold",color="black", angle= 315 , size = 10),
        axis.text.y = element_text(face ="bold",color="black", angle= 0, size = 10)
        )
```

Using the `tidyr` package, the LST Day and Night values for Grand Canyon NP are being gathered in a single column to be used to make a plot including both  `LST_Day_1km` and `LST_Night_1km`.

```{r}
lst_GC_DN <- tidyr::gather(lst_GC, key = Tstat , value = LST, MOD11A2_061_LST_Day_1km, MOD11A2_061_LST_Night_1km)

# Print the first rows of the dataframe
head(lst_GC_DN)
```

Next, plot LST Day and Night as a time series with some additional formatting.

```{r,fig.width = 12, fig.height=5}
ggplot(lst_GC_DN)+
  geom_line(aes(x= Date, y = LST, color = Tstat), size=1)+
  geom_point(aes(x= Date, y = LST, color = Tstat), shape=18 , size = 3)+
  scale_fill_manual(values=c("red", "blue"))+
  scale_color_manual(values=c('red','blue'))+
  labs(title = "Time Series",
       x = "Date",
       y = sprintf( "LST_Day_1km (%s)",unit))+
  scale_x_date(date_breaks = "16 day", limits = as.Date(c('2018-01-01','2019-01-01')))+  
  scale_y_continuous(limits = c(250, 325), breaks = seq(250, 325, 10))+
  theme(plot.title = element_text(face = "bold",size = rel(2.5), hjust = 0.5),
        axis.title = element_text(face = "bold",size = rel(1)),
        panel.background = element_rect(fill = "lightgray", colour = "black"),
        axis.text.x = element_text(face ="bold",color="black", angle= 315 , size = 10),
        axis.text.y = element_text(face ="bold",color="black", angle= 0, size = 10),
        legend.position = "bottom",
        legend.title = element_blank()
        )
```

Finally, bring in the daytime LST data from Zion National Park, and compare with daytime LST at Grand Canyon National Park, shown below in a scatterplot using `ggplot2` package. 
Here, the `dplyr` is used to extract the LST_DAY_1km for Zion National Park.  

```{r}
lst_Z <- df %>%
  # Filter fill value
  filter(MOD11A2_061_LST_Day_1km != fill_value) %>%
  # Filter Zion national park
  filter(Category == "Zion") %>%
  # Select desired columns
  select(Date, MOD11A2_061_LST_Day_1km)
```

Make a scatterplot.

```{r}
ggplot() +
  geom_point(aes(x=lst_Z$MOD11A2_061_LST_Day_1km, y=lst_GC$MOD11A2_061_LST_Day_1km), shape=18 , size = 3, color="blue") +
  labs(title = "MODIS LST: Grand Canyon vs. Zion National Park, 2018",
       x = sprintf("Zion: LST_Day_1km (%s)",unit),
       y = sprintf( "Grand Canyon: LST_Day_1km (%s)",unit)) +
  theme(plot.title = element_text(face = "bold",size = rel(1.5), hjust = 0.5),
        axis.title = element_text(face = "bold",size = rel(1)),
        panel.background = element_rect(fill = "lightgray", colour = "black"),
        axis.text.x = element_text(face ="bold",color="black", size = 10),
        axis.text.y = element_text(face ="bold",color="black", size = 10)
   )
```

This example can provide a template to use for your own research workflows. Leveraging the AppEEARS API for searching, extracting, and formatting analysis ready data, and loading it directly into R means that you can keep your entire research workflow in a single software program, from start to finish.

***

## Contact Information  
 LP DAAC ^1^
 Contact: LPDAAC@usgs.gov   
 Voice: +1-866-573-3222   
 Organization: Land Processes Distributed Active Archive Center (LP DAAC)  
 Website: https://lpdaac.usgs.gov/  
 Date last modified: 10-01-2024      
 
 ^1^ KBR, Inc., contractor to the U.S. Geological Survey, Earth Resources Observation and Science (EROS) Center,  
 Sioux Falls, South Dakota, USA. Work performed under USGS contract G15PD00467 for LP DAAC^2^.  
 
 ^2^ LP DAAC Work performed under NASA contract NNG14HH33I.
